Pool Model - A unified Mathematical\\ Modelling Framework for Food Science

### 2025-11-05 paper/pool_model.tex
reformulate derivation of new equations for ABM
- remove table in favor of list of equations for mappings
- also adapt supplement
- add derivation for transition from lag-phase to growth phase for ABM

        \dot{[R]} &= u \sum\limits_{c_A}\frac{V_{c_A}}{V_D} [R_e] \delta(x-x_{c_A})\\
            &-\frac{u}{1+\nu_I [I]} \sum\limits_{c_B}\frac{V_{c_B}}{V_D} [R_e] \delta(x-x_{c_B})\\
            &+D_R \Delta [R_e]\\

@@

We are thus able to derive the \ac{ode} pool model from its \ac{abm} version via a process of homogenization (see Appendix~\ref{subsec:supplement-abm-details}).

@@

    \dot{R} &= -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
    \dot{I} &= b V\\
    \dot{V} &= \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V.

@@

We compare the most general version and thus omit species $A$ for the moment.
\begin{table*}
    \renewcommand{\arraystretch}{1.5}
    \begin{tabularx}{0.8\textwidth}{@{}l *3{>{\centering\arraybackslash}X}@{}}
        & \textbf{\acs{ode}}
        & \textbf{\acs{abm}}
        & \textbf{Mappings}\\

        \multirow{2}{*}{Resource}
        & \multirow{2}{*}{$\displaystyle \dot{R} = - \frac{\alpha}{(1 + \mu_I I) N_t} R G$}
        & \multirow{2}{*}{$\displaystyle \dot{R} = -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V$}
        & $\displaystyle N_t = [R_0] V_D \sigma$\\

        &&& $\displaystyle \mu_I = \frac{\nu_I}{V_D}$\\

        Volume
        & $\displaystyle \dot{G} = \frac{\alpha}{1 + \mu_I I} R G$
        & $\displaystyle \dot{V} = \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V$
        & $\displaystyle \alpha = u[R_0]\sigma$\\

        Inhibitor
        & $\displaystyle \dot{I} = \kappa G_A$
        & $\displaystyle \dot{I} = b V$
        & $\displaystyle \kappa = b$\\
    \caption{Mappings between \ac{ode} and \ac{abm} parameters.}
    \label{tab:mappings-ode-abm}
\end{table*}
We can use these mappings to numerically compare both approaches.
Table~\ref{tab:spatial_limit_init} shows a comparison of parameters of two simulations for the \ac{ode} and \ac{abm} model which have been mapped onto each other.

@@

@@

    \begin{tikzonimage}[width=0.5\columnwidth]

@@

% ================================================================================================
\newpage
\beginsupplement
\section{Appendix}
% ================================================================================================
\subsection{General mathematical formalism}
\label{ssec:supplement1}
Consider the system consisting of $N$ species in $M$ different states.
In the general representation, all discussed in the main part systems  can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
\begin{equation}
   \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
\label{eq:model_ODE_x}
\end{equation}
where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector describing bacterial count.
In matrix form it can be written as
\begin{equation}
    \mbx = \begin{pmatrix}
        x^1_1  & \dots & x^1_N  \\
        \vdots &       & \vdots \\
        x^M_1  & \dots & x^M_N  \\
            \end{pmatrix}
    \label{eq:model_bact}
\end{equation}
with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
\begin{equation}
    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu).
    \label{eq:model_microenv}
\end{equation}
Here function $\mbH$ is in general case also non-linear.\\
%
The advantage of the following formulation is its generality.
Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.\\
%
To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
\begin{equation}
    N_i = \sum_k \mbx_i^k.
\label{eq:N_pro_species}
\end{equation}
While the total bacterial count of all species of our system is
\begin{equation}
    N = \sum_i \sum_k \mbx_i^k.
\label{eq:N_total}
\end{equation}
%
%
%The three pool system described with equations~\ref{eq:ode_3pools} will be written as
%\begin{equation}
%    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
%        -(\mu + \lambda) & 0             & 0      \\
%        \lambda          & \alpha - \mu' & 0      \\
%        \mu              & \mu'          & -\beta 
%    \end{pmatrix}.
%\label{eq:x_repres_3pools}
%\end{equation}
%
Using such matrix representation of the bacterial system, we can rewrite, for example, the three-pool model with limiting resource described with equations~\ref{eq:ode_3pool_resource} in a more compact form.
The concentrations of the pools are then the components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
%
\begin{align}
    \dot \mbx  &= \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
        -(\mu + \lambda) & 0               & 0      \\
        \lambda          & \alpha R - \mu' & 0      \\
        \mu              & \mu'            & -\beta 
    \end{pmatrix} \mbx\\
    \dot \mbm &= \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
\end{align}
Separating linear and non-linear parts in two terms yields
\begin{equation}
    \dot \mbx = \begin{pmatrix}
        -(\mu + \lambda) & 0       & 0 \\
         \lambda         & \alpha  & 0 \\
         \mu &  \mu'& - \beta 
    \end{pmatrix} 
    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
\label{eq:x_repres_3pools_resource2}
\end{equation}
Here the first term is the linear part and the second term define the quadratic behavior of the system.
%
\paragraph{Generalized Lotka-Volterra}

@@

The simpler version of one state per species (equation~\ref{eq:LV_simple}) can be simply written in a general mathematical form as generalized Lotka-Volterra:
\begin{equation}
    \frac{d x_i}{d t} = x_i f_i (\mbx), \hspace{3mm} \mathbf{f} = \mathbf{r} + A \mbx.
\label{eq:gLV}
\end{equation}

\subsection{\acs{abm} details}
\label{subsec:supplement-abm-details}

\subsubsection*{Deriving the pool model}
Since the \ac{abm} can be considered more fundamental than the \ac{ode} model, we should be able to derive the latter one from the former.
This can be achieved by starting with equations~\eqref{eq:spatial-full-system-simplified}.
We integrate the spatial equations over the total domain starting with the resource concentration $[R]$.
\begin{alignat}{7}
    \int\limits_D\dot{[R]}\text{dVol} 
    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) &&+ D_R\Delta [R]\bigg]\text{dVol}\\
    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_D D_R\Delta [R]\text{dVol}\\
    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_{\partial D} D_R\vec{\nabla} [R]\vec{\text{dA}}\\
    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c \int\limits_D \frac{1}{V_D} [R] \delta(x-x_c)\text{dVol}\\
    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]_{x=x_c}\\
    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]\\
    &= &&&&-\frac{u}{1+\nu_I [I]V_D V_D^{-1}} V [R]\frac{V_D}{V_D}\\
    \dot{R} &= &&&&-\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
\end{alignat}
We can split up the integral into two parts.
The second part can then be transformed into an integral over the flux of the resource along the boundary.
Since we chose Neumann boundary conditions, this contribution vanishes identically.
Afterwards, the first part can be evaluated by using the delta-distribution.
Note that we additionally need to consume a $V_D^{-1}$ term since our volume is finite.
We evaluate the resource concentration $[R]$ at the position of the individual cells and sum over all contributions.
Now we assume homogeneity in the sense, that we can reduce $[R]$ to only a single variable instead of considering its spatial distribution.
This means that every contribution does not depend on the position at which it was evaluated, i.e., $[R]_{x=x_c}=[R]$.
Finally, we can perform the sum to gain the total volume of all cells $V=\sum\limits_c V_c$.
\begin{alignat}{7}
    \int\limits_D\dot{[I]} &= \int\limits_D \bigg[&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) &&+ D_I\Delta [I]\bigg]\text{dVol}\\
    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_D D_I\Delta [I]\text{dVol}\\
    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_{\partial D} D_I\nabla [I]\vec{\text{dA}}\\
    &= &&\sum\limits_c b V_c\int\limits_D \frac{1}{V_D} &&\delta(x-x_c) \text{dVol}\\
    \dot{I} &= b V
\end{alignat}
Following the same principle as before, we can integrate the inhibitor equation.
For the equation describing the growth of the cell volume, we simply add up all cells to obtain the total volume of all cells.
\begin{alignat}{3}
    \sum\limits_c\dot{V_c}
    &= \sum\limits_c&&\frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
    &= &&\frac{u}{1+\nu_I [I]} \sigma [R]V\\
    \dot{V} &= &&\frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V
\end{alignat}
We can thus summarize the equations obtained
\begin{align}
    \dot{R} &= -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
    \dot{I} &= b V\\
    \dot{V} &= \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V.
\end{align}
It is clear that the original model is recovered this way.
To obtain the full equations, we additionally need to consider transition between states of lag-phase and growth phase not depicted here.
The preceding equations all assume that cells are in growth phase.
It is however straightforward, to set up these equations.
%
%
% ................................................................................................
\subsubsection{Full list of parameters}
\label{sec:supplement-full-parameters}

\begin{table}

@@

\subsubsection{Homogeneous and inhomogeneous initial cell distributions}

@@

>>>>

        \dot{[R]} &= u \sum\limits_{c_A}\frac{V_{c_A}}{V_D} [R] \delta(x-x_{c_A})\\
            &-\frac{u}{1+\nu_I [I]} \sum\limits_{c_B}\frac{V_{c_B}}{V_D} [R] \delta(x-x_{c_B})\\
            &+D_R \Delta [R]\\

@@

We are thus able to derive the \ac{ode} pool model from its \ac{abm} version via a process of homogenization (see Appendix~\ref{sec:supplement-abm-derivation}).

@@

    \dot{L_A} &= -\lambda L_A\frac{R}{R_0}\\
    \dot{L_B} &= -\lambda L_B\frac{R}{R_0}\\
    \dot{V_A} &= u\sigma R \frac{V_A}{V_D} + \lambda L_A\frac{R}{R_0}\\
    \dot{V_B} &= \frac{u}{1+\nu_I V_D^{-1} I} \sigma R \frac{V_B}{V_D} + \lambda L_B\frac{R}{R_0}\\
    \dot{R} &= - R \frac{V_A}{V_D} - \frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V_B\\
    \dot{I} &= b V_A

@@

When comparing these equations, we can map them onto each other via the relations
        N_t &= [R_0]V_D\sigma\\
        \mu_I &= \frac{\nu_I}{V_D}\\
        \alpha &= u[R_0]\sigma\\
        \kappa &= \sigma.
    \label{eq:ode-abm-mappings}
Furthermore, we can use these mappings to numerically compare both approaches.
Table~\ref{tab:spatial_limit_init} shows an example comparison of parameters of the \ac{ode} and \ac{abm} which have been mapped onto each other.

@@

\begin{figure*}

@@

    \begin{tikzonimage}[width=\columnwidth]

@@

% ================================================================================================
\newpage
\beginsupplement
% ================================================================================================
\section{General mathematical formalism}
\label{ssec:supplement1}
Consider the system consisting of $N$ species in $M$ different states.
In the general representation, all discussed in the main part systems  can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
\begin{equation}
   \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
\label{eq:model_ODE_x}
\end{equation}
where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector describing bacterial count.
In matrix form it can be written as
\begin{equation}
    \mbx = \begin{pmatrix}
        x^1_1  & \dots & x^1_N  \\
        \vdots &       & \vdots \\
        x^M_1  & \dots & x^M_N  \\
            \end{pmatrix}
    \label{eq:model_bact}
\end{equation}
with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
\begin{equation}
    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu).
    \label{eq:model_microenv}
\end{equation}
Here function $\mbH$ is in general case also non-linear.\\
%
The advantage of the following formulation is its generality.
Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.\\
%
To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
\begin{equation}
    N_i = \sum_k \mbx_i^k.
\label{eq:N_pro_species}
\end{equation}
While the total bacterial count of all species of our system is
\begin{equation}
    N = \sum_i \sum_k \mbx_i^k.
\label{eq:N_total}
\end{equation}
%
%
%The three pool system described with equations~\ref{eq:ode_3pools} will be written as
%\begin{equation}
%    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
%        -(\mu + \lambda) & 0             & 0      \\
%        \lambda          & \alpha - \mu' & 0      \\
%        \mu              & \mu'          & -\beta 
%    \end{pmatrix}.
%\label{eq:x_repres_3pools}
%\end{equation}
%
Using such matrix representation of the bacterial system, we can rewrite, for example, the three-pool model with limiting resource described with equations~\ref{eq:ode_3pool_resource} in a more compact form.
The concentrations of the pools are then the components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
%
\begin{align}
    \dot \mbx  &= \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
        -(\mu + \lambda) & 0               & 0      \\
        \lambda          & \alpha R - \mu' & 0      \\
        \mu              & \mu'            & -\beta 
    \end{pmatrix} \mbx\\
    \dot \mbm &= \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
\end{align}
Separating linear and non-linear parts in two terms yields
\begin{equation}
    \dot \mbx = \begin{pmatrix}
        -(\mu + \lambda) & 0       & 0 \\
         \lambda         & \alpha  & 0 \\
         \mu &  \mu'& - \beta 
    \end{pmatrix} 
    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
\label{eq:x_repres_3pools_resource2}
\end{equation}
Here the first term is the linear part and the second term define the quadratic behavior of the system.
%
\paragraph{Generalized Lotka-Volterra}

@@

The simpler version of one state per species (equation~\ref{eq:LV_simple}) can be simply written in a general mathematical form as generalized Lotka-Volterra:
    \frac{d x_i}{d t} = x_i f_i (\mbx), \hspace{3mm} \mathbf{f} = \mathbf{r} + A \mbx.
\label{eq:gLV}
\vfill
\pagebreak
\section{Deriving the pool model}
\label{sec:supplement-abm-derivation}
\paragraph{Reactions}
Since the \ac{abm} can be considered more fundamental than the \ac{ode} model, we should be able to derive the latter one from the former.
This can be achieved by starting with equations~\eqref{eq:spatial-full-system-simplified}.
We integrate the spatial equations over the total domain starting with the resource concentration $[R]$.
\begin{alignat}{8}
    \int\limits_D\dot{[R]}\text{dVol}
    &= \int\limits_D\bigg[&-&u \sum\limits_{c_A}\frac{V_{c_A}}{V_D} [R] \delta(x-x_{c_A})
        &-&\frac{u}{1+\nu_I [I]} \sum\limits_{c_B}\frac{V_{c_B}}{V_D} [R] \delta(x-x_{c_B})
        +D_R \Delta [R]\bigg]\text{dVol}\\
    &= &-&u\sum\limits_{c_A}V_{c_A}\int\limits_D [R] \frac{\delta(x-x_{c_A})}{V_D} \text{dVol}
        &-&\frac{u}{1+\nu_I [I]} \sum\limits_{c_B}V_{c_B} \int\limits_D [R] \frac{\delta(x-x_{c_B})}{V_D}\text{dVol}
        +D_R \int\limits_{\partial D} \nabla[R]\text{dVol}\\
    &= &-&u\sum\limits_{c_A} V_{c_A} [R]_{x=x_{c_A}}
        &-& \frac{u}{1+\nu_I [I]} \sum\limits_{c_B} V_{c_B} [R]_{x=x_{c_B}}\\
    &= &-&u\sum\limits_{c_A} V_{c_A} [R]
        &-& \frac{u}{1+\nu_I [I]} \sum\limits_{c_B} V_{c_B} [R]\\
    &= &-&uV_A [R] &-&\frac{u}{1+\nu_I[I]} V_B [R]\\
    &= &-&u \frac{V_A}{V_D}R &-&\frac{u}{1+\nu_I V_D^{-1} I} \frac{V_B}{V_D}R
We can split up the integral into three parts.
The last part can then be transformed into an integral over the flux of the resource along the boundary.
Since we chose Neumann boundary conditions, this contribution vanishes identically.
Afterwards, the remaining parts can be evaluated by using the delta-distribution.
Note that we additionally need to consume a factor of $V_D^{-1}$ which is the inverse of the domain
volume when evaluating the delta function.
We evaluate the resource concentration $[R]$ at the position of the individual cells and sum over all contributions.
Now we assume homogeneity in the sense, that we can reduce $[R]$ to only a variable with time-dependence, with homogeneous spatial distribution.
This means that every contribution does not depend on the position at which it was evaluated, i.e., $[R]_{x=x_c}=[R]$.
Finally, we can perform the sum to gain the total volume of all cells $V_{A/B}=\sum\limits_c V_{c_{A/B}}$.
Following the same principle as before, we can integrate the inhibitor equation.
\begin{alignat}{7}
    \int\limits_D\dot{[I]} &= \int\limits_D \bigg[&&\sum\limits_{c_A} b\frac{V_{c_A}}{V_D}&&\delta(x-x_{c_A}) &&+ D_I\Delta [I]\bigg]\text{dVol}\\
    &= \int\limits_D&&\sum\limits_{c_A} b\frac{V_{c_A}}{V_D}&&\delta(x-x_{c_A}) \text{dVol} &&+ \int\limits_D D_I\Delta [I]\text{dVol}\\
    &= \int\limits_D&&\sum\limits_{c_A} b\frac{V_{c_A}}{V_D}&&\delta(x-x_{c_A}) \text{dVol} &&+ \int\limits_{\partial D} D_I\nabla [I]\vec{\text{dA}}\\
    &= &&\sum\limits_{c_A} b V_{c_A}\int\limits_D \frac{1}{V_D} &&\delta(x-x_{c_A}) \text{dVol}
\paragraph{Lag-Phase}
Cells are initially assumed to be in a lag-phase which they leave with probability $P(\text{transition})$.
Dividing this probability by the time increment $\Delta t$, gives us the rate at which the transition is occuring.
This expression can be transfered to a collective rate of change by multiplying with the volume of cells in lag-phase $L_{A/B}$.
    P(\text{transition}) &= \Delta t \lambda \frac{[R]}{[R_0]}\\
    \frac{1}{\Delta t}P(\text{transition}) &= \lambda \frac{[R]}{[R_0]}\\
    \frac{L_{A/B}}{\Delta t}P(\text{transition}) &= \lambda L_{A/B}\frac{R}{R_0}\\
    \partial_t L_{A/B} &= -\lambda L_{A/B}\frac{R}{R_0}
In the last step, we take the expectation value and equate it with the negative of its derivative $\partial_t L_{A/B}$.
\paragraph{Growth}
For the equation describing the growth of the cell volume, we simply add up cells for the respective species to obtain their total volume.
\begin{alignat}{5}
    \sum\limits_{c_A}\dot{V_{c_A}}
    &= &\sum\limits_{c_A} u& \sigma [R]V_{c_A} + \lambda L_A\frac{R}{R_0} &= &&\sigma [R]V_A + \lambda L_A\frac{R}{R_0}\\
    \sum\limits_{c_B}\dot{V_{c_B}}
    &= &\frac{u}{1+\nu_I [I]} &\sigma [R]V_B + \lambda L_B\frac{R}{R_0} &= &\sum\limits_{c_B} \frac{u}{1+\nu_I [I]} &\sigma [R]V_{c_B} + \lambda L_B\frac{R}{R_0}
In addition, we can reformulate $\tilde{\sigma}=\sigma/V_D$ thus redefining the conversion rate $\sigma$ in terms of the total volume $V_D$.
We can thus summarize the equations obtained
    \dot{L_A} &= -\lambda L_A\frac{R}{R_0}\\
    \dot{L_B} &= -\lambda L_B\frac{R}{R_0}\\
    \dot{V_A} &= u\sigma R \frac{V_A}{V_D} + \lambda L_A\frac{R}{R_0}\\
    \dot{V_B} &= \frac{u}{1+\nu_I V_D^{-1} I} \sigma R \frac{V_B}{V_D} + \lambda L_B\frac{R}{R_0}\\
    \dot{R} &= - R \frac{V_A}{V_D} - \frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V_B\\
    \dot{I} &= b V_A
In this way, we are able to recover the \ac{ode} model.
\vfill
\pagebreak
\section{Full list of parameters}
We utilize the calculate mappings (see equations~\eqref{eq:ode-abm-mappings}) between the \ac{ode} formulation and \ac{abm} to compare values for parameters which serve a similar purpose within the two simulations.
Table~\ref{tab:spatial_limit_init} lists these parameters together with their values.
Most parameters have clear $1:1$ mappings to ones within the other model.
However, some such as the ones related to spatial dynamics (damping $\eta$, mass $m$, potential
strength $\epsilon$, diffusion $D_R,D_I$) do not.
\begin{table}[H]

@@

\vfill
\pagebreak
\section{Homogeneous and inhomogeneous initial cell distributions}

@@

\begin{figure}[H]
